Manuscript

### 2022-02-08 analysis/manuscript.Rmd

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)

tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)

@@

1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$).

A summary of the these results can be seen below:

@@

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(description = outcome_description[1], n = n(), 
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
  dplyr::rename(trait_ID = outcome_ID)

confounder_summary <- rbind(non_sig_confounders, sig_confounders)

format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))

DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

@@

# Effect of sex, age and time together on causal effects in couples.

@@

A summary of the $\omega$, $\rho$, $\gamma$ and $\rho_{resid} + \gamma$ estimates are summarized in the table below. Those with BF-significant differences between the estimates (according to a Z-test, p < 0.05/`r dim(data)[1]`) are given as T/F in the last three columns.

@@

>>>>

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)

@@

1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
5. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
6. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.

A summary of the top 5 confounders for each `trait_ID` can be seen below:

@@

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 5) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C_ratio = max(corr_due_to_confounding_ratio), 
                   min_C_ratio = min(corr_due_to_confounding_ratio)) %>%
  dplyr::rename(trait_ID = outcome_ID)

#confounder_summary <- rbind(non_sig_confounders, sig_confounders)

format_round_cols <- colnames(dplyr::select_if(sig_confounders, is.numeric))

DT::datatable(sig_confounders, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

@@

```{r confounder_effect_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plot_data <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 1) %>%
  mutate(corr_minus_MR_est = abs(between_trait_correlation)- abs(IVW_beta))

# plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
#                                                        y = MR_est, 
#                                                        label = Phenotype, 
#                                                        color = factor(diff_p_sig)),) +
#   geom_point(alpha = 1.2) +
#   scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
#   #geom_text(show.legend = FALSE) +
#   geom_abline(slope=1, intercept=0) +
#   
#   ggplot2::geom_errorbar(
#     ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
#     colour = "grey",
#     width = 0
#   ) +
#   ggplot2::geom_errorbar(
#     ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
#     colour = "grey",
#     width = 0
#   ) +
#   
#   theme_minimal() +
#   ggplot2::labs(x = paste0("Couple correlation"),
#                 y = paste("Standardized MR effect")) +
#   theme(legend.position="bottom") + theme(legend.title=element_blank()) +
#   geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
#                   nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
#                   direction = "x")

```
# Effect of sex, age and time together on causal effects in couples.

@@

A summary of the $\omega$, $\rho$, $\gamma$ and $\rho_{resid} + \gamma$ estimates are summarized in the table below. The differences between the estimates (according to a Z-test) are given in the last three columns. Trait pairs were pruned such that:  if there is a pair A-B and C-D and if [max(corr(A,C)*corr(B,D),corr(A,D)*corr(B,C))] > `r corr_trait_threshold` then one pair is dropped. Pairs were prioritized based on the $\omega$ p-value.

@@

tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune)

data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune
data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
data$gam_rho_resid <- data$rho_resid + data$gam_beta
